---
title: "Scraped Old Evals"
output:
  html_document:
    df_print: paged
---

Load libraries and data
```{r}
options(warn=-1)
library(tidyverse)
library(bigrquery)
library(lubridate)
library(statar)
library(knitr)
library(stringi)
library(broom)
library(ggrepel)
library(ggfortify)

setwd("C:/Users/victo/Box/dairy_matching/code/AISS data")

evals <- read.csv("../../data/old_evals.csv")
```

Select and rename columns
```{r}
evals |>
  rename(reg_name = Registration.Name, 
         reg_num = Registration.Number,
         naab_code = Primary.NAAB.Code,
         status = Status,
         controller_name = Controller.Name,
         dob = Date.of.Birth,
         aaa_codes =  aAa,
         genetic_codes = Genetic.Codes..,
         beta_casein = Beta.Casein,
         kappa_casein = Kappa.Casein, 
         beta_lactaglobulin = Beta.Lactoglobulin,
         NM = NM.,
         NM_percentile = NM.Percentile,
         CM = CM.,
         FM = FM.,
         GM = GM.,
         pta_milk = Milk,
         pta_scs = SCS, 
         pta_fat = Fat...,
         pta_protein = Protein...,
         pta_pl = Productive.Life,
         pta_liv = Livability,
         pta_dpr = DPR, 
         pta_hcr = HCR, 
         pta_ccr  = CCR, 
         pta_scr = SCR, 
         pta_gest_length = Gestation.Length, 
         pta_gest_length_obs = Gestation.Length.obs, 
         pta_efcalving = Early.First.Calving,
         pta_efcalving_obs = Early.First.Calving.obs,
         pta_type = PTA.Type, 
         udder_composite = UDC,
         pta_stature = Stature,
         pta_strength = Strength,
         pta_body_depth = Body.Depth,
         pta_dairy_form = Dairy.Form,
         pta_rump_angle = Rump.Angle, 
         pta_thurl_width = Thurl.Width,
         pta_rear_legs_side = Rear.Legs.Side.View,
         pta_foot_angle  = Foot.Angle, 
         pta_fore_udder = F.U..Attachment,
         pta_rear_udder_height = Rear.Udder.Height,
         pta_rear_udder_width = Rear.Udder.Width,
         pta_udder_cleft = Udder.Cleft,
         pta_udder_depth  = Udder.Depth, 
         pta_teat_front_place = Front.Teat.Placement,
         pta_teat_length = Teat.Length,
         pta_milking_speed = Milking.Speed,
         pta_mastitis = Mastitis,
         pta_metritis = Metritis, 
         pta_disp_abomasum = D..Abom,
         pta_ketosis = Ketosis,
         pta_r_placenta = R..Placenta,
         pta_milk_fever  = Milk.Fever,
         sire_ce = Sire.CE,
         daughter_ce = Daughter.CE,
         sire_ce_obs = Sire.CE.obs,
         daughter_ce_obs = Daughter.CE.obs,
         pta_mobility = Mobility,
         sire_name = Sire,
         mgs_name = MGS, 
         pta_rear_legs_rear = Rear.Legs.Rear.View,
         naab_code_secondary = Secondary.Code.s.,
         haplotypes = Haplotypes,
         feed_saved = Feed.Saved,
         pta_heifer_liv = Heifer.Livability,
         sire_sb = Sire.SB, 
         daughter_sb = Daughter.SB,
         sire_sb_obs = Sire.SB.obs,
         daughter_sb_obs = Daughter.SB.obs,
         pta_foot_leg_score = Feet...Legs.Score,
         pta_teat_rear_place = Rear.Teat.Placement) ->
  evals
```

Parse `production_obs` into number of daughters, herds and percent of US daughters.
```{r}
evals |>
  mutate(num_dtrs = str_split_fixed(production_obs, " ", 9)[,1] |> as.numeric(),
         num_herds = str_split_fixed(production_obs, " ", 9)[,3] |> as.numeric(),
         us_perc = str_split_fixed(production_obs, " ", 9)[,5] |> as.numeric(),
         reliability = str_split_fixed(production_obs, " ", 9)[,8]) |>
  select(-production_obs) |>
  separate(reliability, into = c("reliability", "sign"), sep = -1) |>
  select(-sign) |>
  mutate(us_perc = us_perc/100,
         reliability = as.numeric(reliability)/100) ->
  evals
```

Parse Date of birth
```{r}
evals |>
  mutate(dob = mdy(dob),
         yob = year(dob)) ->
  evals
```

Parse controller names
```{r}
evals |> 
  extract(controller_name, c("controller_name", "controller_number"), 
          regex = "(.+)\\s\\((.+)\\)") |>
  mutate(controller_number = as.numeric(controller_number)) ->
  evals
```

DMS and GFI
```{r}
evals |>
  rename(temp = GFI) |>
  mutate(GFI = ifelse(temp != "-", temp, NA_real_), 
         DMS = as.numeric(DMS)) |>
  separate(GFI, into = c("GFI", "perc"), sep = -1) |>
  mutate(GFI = as.numeric(GFI)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Recode Status
```{r}
evals |>
  rename(status_alt = status) |>
  mutate(status = case_when(
    status_alt == "Active AI Sire" ~ "A",
    status_alt == "Foreign" ~ "F",
    status_alt == "Genomically Tested" ~ "G",
    status_alt == "Inactive" ~ "I",
    status_alt == "Natural Service Bull" ~ "N",
    status_alt == "Progeny Test" ~ "P",
    status_alt == "Unknown" & yob < 2011 ~ "I",
    .default = NA_character_)) ->
  evals
```

Net Merit
```{r}
evals |> 
  separate(NM, into = c("NM", "NM_rel"), sep = ",") |>
  separate(NM_rel, into = c("NM_rel", "chars"), sep = 3) |>
  select(-chars) |>
  mutate(NM = as.numeric(NM), 
         NM_rel = as.numeric(NM_rel)/100) ->
  evals
```
Net Merit Percentile
```{r}
evals |>
  separate(NM_percentile, into = c("NM_percentile", "perc"), sep = -1) |>
  mutate(NM_percentile = as.numeric(NM_percentile)) |>
  select(-perc) ->
  evals
```
PTI, CM, FM and GM
```{r}
evals |>
  mutate(PTI = as.numeric(PTI),
         CM = as.numeric(CM),
         FM = as.numeric(FM),
         GM = as.numeric(GM)) ->
  evals
```

Somatic Cell Score
```{r}
evals |> 
  separate(pta_scs, into = c("pta_scs", "pta_scs_rel"), sep = ",") |>
  separate(pta_scs_rel, into = c("pta_scs_rel", "chars"), sep = -3) |>
  mutate(pta_scs = as.numeric(pta_scs),
         pta_scs_rel = as.numeric(pta_scs_rel)) |>
  select(-chars)  ->
  evals
```
Milk
```{r}
evals |> 
  mutate(pta_milk = as.numeric(pta_milk)) ->
  evals
```

Fat (lbs)
```{r}
evals |> 
  separate(pta_fat, into = c("pta_fat_lb", "pta_fat_lb_rel"), sep = ",") |>
  separate(pta_fat_lb_rel, into = c("pta_fat_lb_rel", "chars"), sep = -2) |>
  mutate(pta_fat_lb = as.numeric(pta_fat_lb),
         pta_fat_lb_rel = as.numeric(pta_fat_lb_rel)) |>
  select(-chars)  ->
  evals
```
Protein (lbs)
```{r}
evals |> 
  separate(pta_protein, into = c("pta_protein_lb", "pta_protein_lb_rel"), sep = ",") |>
  separate(pta_protein_lb_rel, into = c("pta_protein_lb_rel", "chars"), sep = -2) |>
  mutate(pta_protein_lb = as.numeric(pta_protein_lb),
         pta_protein_lb_rel = as.numeric(pta_protein_lb_rel)) |>
  select(-chars)  ->
  evals
```
Productive Life
```{r}
evals |> 
  rename(temp = pta_pl) |>
  mutate(pta_pl = str_split_fixed(temp, " ", 2)[,1],
         pta_pl_rel = str_split_fixed(temp, " ", 2)[,2]) |>
  separate(pta_pl_rel, into = c("pta_pl_rel", "chars"), sep = -3) |>
  mutate(pta_pl = as.numeric(pta_pl),
         pta_pl_rel = as.numeric(pta_pl_rel)/100) |>
  select(-c(chars, temp))  ->
  evals
```

Livability
```{r}
evals |> 
  rename(temp = pta_liv) |>
  mutate(pta_liv = str_split_fixed(temp, " ", 2)[,1],
         pta_liv_rel = str_split_fixed(temp, " ", 2)[,2]) |>
  separate(pta_liv_rel, into = c("pta_liv_rel", "chars"), sep = -3) |>
  mutate(pta_liv = as.numeric(pta_liv),
         pta_liv_rel = as.numeric(pta_liv_rel)/100) |>
  select(-c(chars, temp))  ->
  evals
```
Daughter Pregnancy Rate
```{r}
evals |> 
  rename(temp = pta_dpr) |>
  mutate(pta_dpr = str_split_fixed(temp, " ", 2)[,1],
         pta_dpr_rel = str_split_fixed(temp, " ", 2)[,2]) |>
  separate(pta_dpr_rel, into = c("pta_dpr_rel", "chars"), sep = -3) |>
  mutate(pta_dpr = as.numeric(pta_dpr),
         pta_dpr_rel = as.numeric(pta_dpr_rel)/100) |>
  select(-c(chars, temp))  ->
  evals
```
Heifer Conception rate
```{r}
evals |> 
  rename(temp = pta_hcr) |>
  mutate(pta_hcr = str_split_fixed(temp, " ", 2)[,1],
         pta_hcr_rel = str_split_fixed(temp, " ", 2)[,2]) |>
  separate(pta_hcr_rel, into = c("pta_hcr_rel", "chars"), sep = -3) |>
  mutate(pta_hcr = as.numeric(pta_hcr),
         pta_hcr_rel = as.numeric(pta_hcr_rel)/100) |>
  select(-c(chars, temp))  ->
  evals
```
Cow Conception Rate
```{r}
evals |> 
  rename(temp = pta_ccr) |>
  mutate(pta_ccr = str_split_fixed(temp, " ", 2)[,1],
         pta_ccr_rel = str_split_fixed(temp, " ", 2)[,2]) |>
  separate(pta_ccr_rel, into = c("pta_ccr_rel", "chars"), sep = -3) |>
  mutate(pta_ccr = as.numeric(pta_ccr),
         pta_ccr_rel = as.numeric(pta_ccr_rel)/100) |>
  select(-c(chars, temp))  ->
  evals
```
Sire Conception Rate
```{r}
evals |> 
  rename(temp = pta_scr) |>
  mutate(pta_scr = str_split_fixed(temp, " ", 2)[,1],
         pta_scr_rel = str_split_fixed(temp, " ", 2)[,2]) |>
  separate(pta_scr_rel, into = c("pta_scr_rel", "chars"), sep = -3) |>
  mutate(pta_scr = as.numeric(pta_scr),
         pta_scr_rel = as.numeric(pta_scr_rel)/100) |>
  select(-c(chars, temp))  ->
  evals
```

Gestation length
```{r}
evals |> 
  rename(temp = pta_gest_length) |>
  mutate(pta_gest_length = str_split_fixed(temp, " ", 2)[,1],
         pta_gest_length_rel = str_split_fixed(temp, " ", 2)[,2]) |>
  separate(pta_gest_length_rel, into = c("pta_gest_length_rel", "chars"), sep = -3) |>
  mutate(pta_gest_length = as.numeric(pta_gest_length),
         pta_gest_length_rel = as.numeric(pta_gest_length_rel)/100) |>
  select(-c(chars, temp))  ->
  evals
```
Gestation length Observations
```{r}
evals |>
  rename(temp = pta_gest_length_obs) |>
  mutate(pta_gest_length_obs = str_split_fixed(temp, " ", 4)[,1] |> as.numeric(),
         pta_gest_length_dtrs = str_split_fixed(temp, " ", 4)[,3]|> as.numeric()) |>
  select(-temp) ->
  evals
```

Early First Calving
```{r}
evals |> 
  rename(temp = pta_efcalving) |>
  mutate(pta_efcalving = str_split_fixed(temp, " ", 2)[,1],
         pta_efcalving_rel = str_split_fixed(temp, " ", 2)[,2]) |>
  separate(pta_efcalving_rel, into = c("pta_efcalving_rel", "chars"), sep = -3) |>
  mutate(pta_efcalving = as.numeric(pta_efcalving),
         pta_efcalving_rel = as.numeric(pta_efcalving_rel)/100) |>
  select(-c(chars, temp))  ->
  evals
```
Early First Calving Observations
```{r}
evals |> 
  rename(temp = pta_efcalving_obs) |>
  separate(temp, into = c("efc_obs", "efc_dtrs"), sep = ",") |>
  mutate(pta_efcalving_obs = str_split_fixed(efc_obs, " ", 2)[,1] |> as.numeric(),
         pta_efcalving_dtrs = str_split_fixed(efc_dtrs, " ", 3)[,2] |> as.numeric()) |>
  select(-c(efc_obs, efc_dtrs)) ->
  evals
```
Type elements
```{r}
evals |>
  mutate(pta_type_dtrs = str_split_fixed(type_obs, " ", 6)[, 1] |> as.numeric(),
         pta_type_herds = str_split_fixed(type_obs, " ", 6)[, 3] |> as.numeric(),
         pta_type_rel = str_split_fixed(type_obs, " ", 6)[, 5]) |>
  separate(pta_type_rel, into = c("pta_type_rel", "perc"), sep = -1) |>
  mutate(pta_type_rel = as.numeric(pta_type_rel)/100) |>
  select(-c(type_obs, perc)) ->
  evals
```

Milking Speed
```{r}
evals |>
  separate(pta_milking_speed, into = c("pta_milk_speed", "pta_milk_speed_rel"), sep = ",") |>
  separate(pta_milk_speed_rel, into = c("pta_milk_speed_rel", "temp"), sep = -3) |>
  mutate(pta_milk_speed = as.numeric(pta_milk_speed),
         pta_milk_speed_rel = as.numeric(pta_milk_speed_rel)) |>
  select(-temp) ->
  evals
```
Mastitis
```{r}
evals |>
  rename(temp = pta_mastitis) |>
  mutate(pta_mastitis = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         pta_mastitis_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(pta_mastitis_rel, into = c("pta_mastitis_rel", "perc"), sep = -1) |>
  mutate(pta_mastitis_rel = as.numeric(pta_mastitis_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Metritis
```{r}
evals |>
  rename(temp = pta_metritis) |>
  mutate(pta_metritis = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         pta_metritis_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(pta_metritis_rel, into = c("pta_metritis_rel", "perc"), sep = -1) |>
  mutate(pta_metritis_rel = as.numeric(pta_metritis_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Displaced Abomasum
```{r}
evals |>
  rename(temp = pta_disp_abomasum) |>
  mutate(pta_disp_abomasum = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         pta_disp_abomasum_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(pta_disp_abomasum_rel, into = c("pta_disp_abomasum_rel", "perc"), sep = -1) |>
  mutate(pta_disp_abomasum_rel = as.numeric(pta_disp_abomasum_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Ketosis
```{r}
evals |>
  rename(temp = pta_ketosis) |>
  mutate(pta_ketosis = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         pta_ketosis_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(pta_ketosis_rel, into = c("pta_ketosis_rel", "perc"), sep = -1) |>
  mutate(pta_ketosis_rel = as.numeric(pta_ketosis_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
GPPR
```{r}
evals |>
  mutate(GPPR = as.numeric(GPPR)) ->
  evals
```
GTPI
```{r}
evals |>
  mutate(GTPI = as.numeric(GTPI)) ->
  evals
```
Placenta
```{r}
evals |>
  rename(temp = pta_r_placenta) |>
  mutate(pta_r_placenta = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         pta_r_placenta_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(pta_r_placenta_rel, into = c("pta_r_placenta_rel", "perc"), sep = -1) |>
  mutate(pta_r_placenta_rel = as.numeric(pta_r_placenta_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Milk Fever
```{r}
evals |>
  rename(temp = pta_milk_fever) |>
  mutate(pta_milk_fever = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         pta_milk_fever_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(pta_milk_fever_rel, into = c("pta_milk_fever_rel", "perc"), sep = -1) |>
  mutate(pta_milk_fever_rel = as.numeric(pta_milk_fever_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Sire Calving Ease
```{r}
evals |>
  rename(temp = sire_ce) |>
  mutate(sire_ce = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         sire_ce_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(sire_ce_rel, into = c("sire_ce_rel", "perc"), sep = -1) |>
  mutate(sire_ce_rel = as.numeric(sire_ce_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Daughter Calving Ease
```{r}
evals |>
  rename(temp = daughter_ce) |>
  mutate(daughter_ce = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         daughter_ce_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(daughter_ce_rel, into = c("daughter_ce_rel", "perc"), sep = -1) |>
  mutate(daughter_ce_rel = as.numeric(daughter_ce_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Calving Ease Observations
```{r}
evals |>
  mutate(sire_ce_obs = str_split_fixed(sire_ce_obs, " ", 2)[,1] |> as.numeric(),
         daughter_ce_obs = str_split_fixed(daughter_ce_obs, " ", 2)[,1] |> as.numeric()) ->
  evals
```
Feed Saved
```{r}
evals |>
  rename(temp = feed_saved) |>
  mutate(feed_saved = str_split_fixed(temp, " ", 4)[,1] |> as.numeric(),
         feed_saved_rel = str_split_fixed(temp, " ", 4)[,3]) |>
  separate(feed_saved_rel, into = c("feed_saved_rel", "perc"), sep = -1) |>
  mutate(feed_saved_rel = as.numeric(feed_saved_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Heifer Livability
```{r}
evals |>
  rename(temp = pta_heifer_liv) |>
  mutate(pta_heifer_liv = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         pta_heifer_liv_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(pta_heifer_liv_rel, into = c("pta_heifer_liv_rel", "perc"), sep = -1) |>
  mutate(pta_heifer_liv_rel = as.numeric(pta_heifer_liv_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Daughter SB
```{r}
evals |>
  rename(temp = daughter_sb) |>
  mutate(daughter_sb = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         daughter_sb_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(daughter_sb_rel, into = c("daughter_sb_rel", "perc"), sep = -1) |>
  mutate(daughter_sb_rel = as.numeric(daughter_sb_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Sire SB
```{r}
evals |>
  rename(temp = sire_sb) |>
  mutate(sire_sb = str_split_fixed(temp, " ", 3)[,1] |> as.numeric(),
         sire_sb_rel = str_split_fixed(temp, " ", 3)[,2]) |>
  separate(sire_sb_rel, into = c("sire_sb_rel", "perc"), sep = -1) |>
  mutate(sire_sb_rel = as.numeric(sire_sb_rel)/100) |>
  select(-c(temp, perc)) ->
  evals
```
Daughter Observations
```{r}
evals |>
  rename(temp = daughter_sb) |>
  mutate(daughter_sb = str_split_fixed(temp, " ", 3)[,1] |> as.numeric()) |>
  select(-temp) ->
  evals
```

Sire SB Observations
```{r}
evals |>
  rename(temp = sire_sb_obs) |>
  mutate(sire_sb_obs = str_split_fixed(temp, " ", 3)[,1] |> as.numeric()) |>
  select(-temp) ->
  evals
```

Daughter SB Observations
```{r}
evals |>
  rename(temp = daughter_sb_obs) |>
  mutate(daughter_sb_obs = str_split_fixed(temp, " ", 3)[,1] |> as.numeric()) |>
  select(-temp) ->
  evals
```

GJPI and PPR
```{r}
evals |>
  mutate(GJPI = as.numeric(GJPI),
         PPR = as.numeric(PPR)) ->
  evals
```
Rear Legs
```{r}
evals |>
  mutate(pta_rear_legs_rear = as.numeric(pta_rear_legs_rear)) ->
  evals
```

aAa codes
```{r}
evals |>
  mutate(aaa_codes = as.numeric(aaa_codes)) ->
  evals
```
Remove `reg_num` and split `reg_id` into country and breed
```{r}
evals |>
  mutate(breed = substring(evals$reg_id, 1, 2),
         country = substring(evals$reg_id, 3, 5)) |>
  select(-reg_num) ->
  evals
```

Export results
```{r}
col_order <- c("reg_id", "reg_name", "country", "breed", "naab_code", 
               "naab_code_secondary", "status", "controller_name", 
               "controller_number", "dob", "yob", "aaa_codes",
               "genetic_codes", "beta_casein", "kappa_casein", "beta_lactaglobulin",
               "haplotypes",
               "num_dtrs", "num_herds", "us_perc", "reliability",
               "pta_milk", "pta_fat_lb", "pta_fat_lb_rel", "pta_protein_lb", 
               "pta_protein_lb_rel", "pta_scs", "pta_scs_rel", "pta_pl", "pta_pl_rel",
               "pta_dpr", "pta_dpr_rel", "pta_hcr", "pta_hcr_rel", "pta_ccr", "pta_ccr_rel",
               "pta_scr", "pta_scr_rel", "pta_liv", "pta_liv_rel", "pta_type",
               "pta_type_rel", "pta_type_dtrs", "pta_type_herds",
               "pta_gest_length", "pta_gest_length_rel", "pta_gest_length_obs",
               "pta_gest_length_dtrs", "pta_heifer_liv", "pta_heifer_liv_rel", 
               "pta_efcalving", "pta_efcalving_rel", "pta_efcalving_obs",
               "pta_efcalving_dtrs", "pta_mastitis", "pta_mastitis_rel",
               "pta_metritis", "pta_metritis_rel", "pta_disp_abomasum", "pta_disp_abomasum_rel",
               "pta_ketosis", "pta_ketosis_rel", "pta_r_placenta", "pta_r_placenta_rel",
               "pta_rear_legs_rear", "pta_foot_leg_score", "pta_teat_rear_place",
               "pta_rump_angle", "pta_thurl_width",  "pta_rear_legs_side", 
               "pta_foot_angle", "pta_fore_udder", "pta_rear_udder_height", 
               "pta_rear_udder_width", "pta_udder_cleft", "pta_udder_depth",
               "pta_teat_front_place", "pta_teat_length",
               "pta_milk_fever", "pta_milk_fever_rel", "pta_stature", 
               "pta_strength", "pta_mobility", "pta_milk_speed", "pta_milk_speed_rel",
               "pta_dairy_form", "pta_body_depth",
               "sire_ce", "sire_ce_rel", "sire_ce_obs", 
               "daughter_ce", "daughter_ce_rel", "daughter_ce_obs", "sire_sb", 
               "sire_sb_rel", "sire_sb_obs","daughter_sb", "daughter_sb_rel", 
               "daughter_sb_obs", "udder_composite", "feed_saved",
               "feed_saved_rel", "NM", "NM_rel", "NM_percentile", "PTI", "JUI",
               "CM", "FM", "GM", "DMS", "GTPI", "FLC", "GJPI", "PPR", "GFI", 
               "BWC", "GPPR", "sire_name", "mgs_name", "status_alt")

evals <- evals[, col_order]
```

Save data 
```{r}
write_csv(evals, file = "../../data/old_evals_clean.csv")
```

