---
title: "Status Changes in NAAB Data"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Load libraries
```{r}
options(warn=-1)
library(tidyverse)
library(bigrquery)
library(lubridate)
library(statar)
library(knitr)
library(stringi)
library(broom)
library(ggrepel)
library(ggfortify)

setwd("C:/Users/victo/Box/dairy_matching/code/AISS data")
```

# NAAB Info dataset
```{r}
projectid <- "dairy-168114"
query <- "
SELECT
    reg_id,
    id_number,
    breed,
    country,
    primary_naab_code,
    naab_codes,
    reg_name,
    birthdate,
    semen_release_date,
    genotype_information,
    sire_breed,
    sire_country,
    sire_id_number,
    dam_breed,
    dam_country,
    dam_id_number,
    mgs_breed,
    mgs_country,
    mgs_id_number,
    status
FROM
    CDCB_data.NAAB_Info
"

naab <- bq_project_query(projectid, query)
naab <- bq_table_download(naab)

naab <- naab %>%
  mutate(birthdate = mdy(birthdate),
         yob = year(birthdate))

naab <- naab |>
  mutate(id_number = stringr::str_replace(id_number, "\\.", ""),
         id_number = str_pad(id_number, width = 12, side = "left", pad = "0"),
         reg_id = paste0(breed, country, id_number))
```

# AISS data
```{r}
query <- "SELECT
  country,
  id_number,
  name,
  primary_stud_code,
  sire_id,
  mgs_id,
  period,
  status,
  breed,
  naab_code,
  pta_milk_unadj as pta_milk,
  pta_fat_lb_unadj as pta_fat_lb,
  pta_protein_lb_unadj as pta_protein_lb,
  pta_scs_unadj as pta_scs,
  pta_pl_unadj as pta_pl,
  pta_dpr_unadj as pta_dpr,
  pta_hcr_unadj as pta_hcr,
  pta_ccr_unadj as pta_ccr,
  pta_liv_unadj as pta_liv,
  pta_ssb_unadj as pta_ssb,
  pta_dsb_unadj as pta_dsb,
  pta_type_unadj as pta_type,
  pta_stature_unadj as pta_stature,
  pta_strength_unadj as pta_strength,
  pta_body_depth_unadj as pta_body_depth,
  pta_dairy_form_unadj as pta_dairy_form,
  pta_rump_angle_unadj as pta_rump_angle,
  pta_thurl_width_unadj as pta_thurl_width,
  pta_rear_legs_side_unadj as pta_rear_legs_side,
  pta_rear_legs_rear_unadj as pta_rear_legs_rear,
  pta_foot_angle_unadj as pta_foot_angle,
  pta_foot_leg_score_unadj as pta_foot_leg_score,
  pta_fore_udder_unadj as pta_fore_udder,
  pta_rear_udder_height_unadj as pta_rear_udder_height,
  pta_rear_udder_width_unadj as pta_rear_udder_width,
  pta_udder_cleft_unadj as pta_udder_cleft,
  pta_udder_depth_unadj as pta_udder_depth,
  pta_teat_front_place_unadj as pta_teat_front_place,
  pta_teat_rear_place_unadj as pta_teat_rear_place,
  pta_teat_length_unadj as pta_teat_length,
  pta_mobility,
  pta_gest_length_unadj as pta_gest_length,
  pta_ketosis_unadj as pta_ketosis,
  pta_metritis_unadj as pta_metritis,
  pta_mastitis_unadj as pta_mastitis,
  pta_milk_fever_unadj as pta_milk_fever,
  FM_unadj as FM,
  CM_unadj as CM,
  GM_unadj as GM,
  NM_unadj as NM,
  TPI,
  udder_composite, 
  body_weight_composite,
  feet_legs_composite,
  aaa_rating,
  daughts_in_milk_pta,
  dob,
  semen_price
FROM
  CDCB_data.NAAB_AISS
"

naab_aiss <- bq_project_query(projectid, query)
naab_aiss <- bq_table_download(naab_aiss)

naab_aiss <- naab_aiss %>%
  separate(period, into = c("year", "per_num"), sep = "-", remove = FALSE) |>
  mutate(dob = ymd(dob),
         yob = year(dob))

naab_aiss <-  naab_aiss |>
  mutate(id_number = stringr::str_replace(id_number, "\\.", ""),
         id_number = str_pad(id_number, width = 12, side = "left", pad = "0"),
         reg_id = paste0(breed, country, id_number))
```

Create age and current year variables
```{r}
naab_aiss |> 
  mutate(period = ym(period),
         year = year(period),
         age = year-yob) ->
  naab_aiss
```

Fill out missing values with last missing status
```{r}
naab_aiss|> 
  group_by(reg_id) |> 
  fill(status, .direction = "down") |>
  ungroup() ->
  naab_aiss
```

Status by year of proof
```{r}
naab_aiss |>
  mutate(year_proof = year(period)) ->
  naab_aiss
```

Genomic-proven bulls appear for the first time in 2010.
```{r}
naab_aiss |>
  tab(status, year_proof) |>
  select(c(status, year_proof, `Freq.`)) |> 
  pivot_wider(id_cols = year_proof, names_from = "status", values_from = "Freq.") |>
  data.frame()
```
Characterize status transitions. One way to do so is by creating a dummy variable that tells us whether $status_{t+1} \neq status_{t}$ or not.
```{r}
naab_aiss |>
  filter(!is.na(status)) |>
  arrange(period) |>
  group_by(reg_id) |>
  mutate(status = factor(status),
         obs_num = 1:n(),
         status_1 = dplyr::lead(status, n = 1, order_by = obs_num),
         status_change = ifelse(status != status_1, 1, 0)) |>
  ungroup() |>
  arrange(reg_id, period) ->
  naab_aiss

naab_aiss |>
  tab(status_change)
```
Remove all values with `status_1` missing (last reported period for each bull).
```{r}
naab_aiss |>
  filter(!is.na(status_1)) |>
  tab(status, status_1) |>
  select(c(status, status_1, `Freq.`)) |>
  pivot_wider(id_cols = status, names_from = "status_1", values_from = "Freq.")
```
We have a total of 4725 status changes, 2763 (58%) are from Genomic to Active and 884 (19%) are from genomic to foreign. How many bulls are there actually?
```{r}
naab_aiss |>
  filter(status_change == 1) |>
  group_by(reg_id) |> 
  summarise(num_changes = n()) |> 
  arrange(desc(num_changes)) ->
  changes
```
Out of 4725 changes, we have 4296 unique animals, out of 22149 animals (20%). 3956 (92%) Change of status only once, on the other hand, 7 bulls changed status 5 times.

```{r}
changes |>
  tab(num_changes)
```
```{r}
changes |> 
  filter(num_changes > 1) |>
  select(reg_id) ->
  repeated_ids
```

Characterizing repeated bulls
```{r}
naab_repeated <- naab_aiss |>
  filter(reg_id %in% repeated_ids$reg_id)
naab_repeated <- left_join(naab_repeated, changes, by = "reg_id")

naab_repeated |>
  distinct(reg_id, .keep_all = TRUE) |>
  tab(country)
```
The bulk of changes appear to be from "active to "foreign" (and reverse) 


```{r}
naab_repeated |>
  filter(num_changes == 5 & !is.na(status_1)) |>
  tab(status, status_1)
```
```{r}
naab_repeated |>
  filter(num_changes == 4 & !is.na(status_1)) |>
  tab(status, status_1)
```
```{r}
naab_repeated |>
  filter(num_changes == 3 & !is.na(status_1)) |>
  tab(status, status_1)
```
# Marketing codes
```{r}
query <- "SELECT
  market_code,
  company_name,
  larger_company,
  country,
  assigned_date,
  notes
FROM
  `dairy-168114.DHI_data.company_info`"

companies <- bq_project_query(projectid, query)
companies <- bq_table_download(companies)
```

```{r}
companies |>
  mutate(parent_company = case_when(
    str_detect(companies$company_name, "^.*EVOLUTION|Evolution.*$") == TRUE ~ "Synetics",
    str_detect(companies$company_name, "^.*Genex.*$") == TRUE ~ "Genex",
    str_detect(companies$company_name, "^.*Select Sires|Select Sire|Advanced Genetics|Sire Power|Cache Valley|
               All West|World Wide Sires|Genervations.*$") == TRUE ~ "Select Sires",
    str_detect(companies$company_name, "^.*ABS|South Texas.*$") == TRUE ~ "ABS Global",
    str_detect(companies$company_name, "^.*Trans Ova|URUS.*$") == TRUE ~ "Urus",
    str_detect(companies$company_name, "^.*STgenetics|Inguran|TWG|605|Sexing Technologies|ST-Ireland|ST-France|ST Germany GmbH|
               ST-New Zealand|ST-New Zealand|Taurus Service|Trans-America|Inguran.*$") == TRUE ~ "STgenetics",
    str_detect(companies$company_name, "^.*Semex Alliance|Semex|Special Sires.*$") == TRUE ~ "Semex Alliance",
    str_detect(companies$company_name, "^.*Alta Genetics.*$") == TRUE ~ "Alta Genetics",
    str_detect(companies$company_name, "^.*Vogler.*$") == TRUE ~ "Vogler",
    str_detect(companies$company_name, "^.*Total Livestock.*$") == TRUE ~ "Total Livestock Genetics, Pty Ltd.",
    str_detect(companies$company_name, "^.*CRV.*$") == TRUE ~ "CRV Holding",
    str_detect(companies$company_name, "^.*Jetstream.*$") == TRUE ~ "Jetstream Genetics",
    str_detect(companies$company_name, "^.*Blondin Sires, Inc.*$") == TRUE ~ "Blondin Sires, Inc.",
    str_detect(companies$company_name, "^.*Geno AI Center.*$") == TRUE ~ "Geno AI Center",
    str_detect(companies$company_name, "^.*Hawkeye Breeders.*$") == TRUE ~ "Hawkeye Breeders",
    str_detect(companies$company_name, "^.*Viking Genetics.*$") == TRUE ~ "Viking Genetics",
    str_detect(companies$company_name, "^.*Elgin Breeding.*$") == TRUE ~ "Elgin Breeding Service",
    str_detect(companies$company_name, "^.*JLG Enterprises|Glaz-Way.*$") == TRUE ~ "JLG Enterprises",
    str_detect(companies$company_name, "^.*North Florida Holsteins.*$") == TRUE ~ "North Florida Holsteins",
    str_detect(companies$company_name, "^.*Global Ag Alliance.*$") == TRUE ~ "Global Ag Alliance, Inc.",
    str_detect(companies$company_name, "^.*Accelerated Genetics.*$") == TRUE ~ "Accelerated Genetics",
    str_detect(companies$company_name, "^.*International Protein Sires.*$") == TRUE ~ "International Protein Sires",
    str_detect(companies$company_name, "^.*Cooperative Resources.*$") == TRUE ~ "CRI",
    str_detect(companies$company_name, "^.*German Genetics.*$") == TRUE ~ "German Genetics International (GGI)")) -> 
  companies
```
Total Livestock Genetics, Pty Ltd.
```{r}
companies |>
  mutate(assigned_date = ymd_hms(assigned_date),
         parent_company = ifelse(!is.na(parent_company), parent_company, company_name)) ->
  companies
```

```{r}
naab_aiss <- left_join(naab_aiss, companies, 
                       by = join_by(primary_stud_code == market_code))
```

Characterizing `G` to `A`
```{r}
naab_aiss |>
  filter(status == "G" & status_1 == "A") |>
  select(reg_id) ->
  movers

naab_aiss |>
  filter(reg_id %in% movers$reg_id) ->
  naab_movers
```


```{r}
naab_movers |>
  group_by(reg_id) |>
  arrange(period) |>
  slice_head(n = 1) |>
  ungroup() |>
  tab(parent_company)
```
```{r}
naab_movers |>
  group_by(reg_id) |>
  arrange(period) |>
  slice_head(n = 1) |>
  ungroup() |>
  tab(breed)
```


